---
title: "hw03-xiaoya-li.Rmd"
author: "Xiaoya Li"
date: "October 6, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)

teams <- read.csv("../../data/nba2017-teams.csv")
```

# Ranking of Teams

### Basic Rankings

**Start by ranking the teams according to salary, arranged in decreasing order.The vertical red line
is the average team salary.**

```{r}
ggplot(data = teams, aes(x = reorder(team, salary), y = salary)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(yintercept = mean(teams$salary), size = 2, colour = rgb(1,0.3,0,0.5)) +
  labs(y = "Salary (in millions)", x = "Team", title = "NBA Teams ranked by Total Salary")
```


**Create another bar chart of teams ranked by total points. The vertical red line is the average
team points.**

```{r}
ggplot(data = teams, aes(x = reorder(team, points), y = points)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(yintercept = mean(teams$points), size = 2, colour = rgb(1,0.3,0,0.5)) +
  labs(y = "Total Points", x = "Team", title = "NBA Teams ranked by Total Points")
```

**Use efficiency to obtain a third kind of ranking, and create an associated bar chart of
teams ranked by total efficiency. The vertical red line is the average team efficiency.**

```{r}
ggplot(data = teams, aes(x = reorder(team, efficiency), y = efficiency)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(yintercept = mean(teams$efficiency), size = 2, colour = rgb(1,0.3,0,0.5)) +
  labs(y = "Total Efficiency", x = "Team", title = "NBA Teams ranked by Total Efficiency")
```

**Provide concise descriptions of the obtained rankings so far.**\
- *ranked by Total Salary: Around half of the teams are above the average salary, where CLE ranked the first with more than 120 millions total salary.*\
- *ranked by Total Points: More than half of the teams are above the average total points, where GSW ranked the first with more than 8750 total points.*\
- *ranked by Total Efficiency: Only 9 teams out of 30 are above the average efficiency, where CLE ranked the first with more than 175 index of efficiency.*\

# Principal Components Analysis (PCA)

**Perform a principal components analysis (PCA) on the following variables, to use the first
principal component (PC1) as another index to rank the teams:**

```{r}
teams_pca <- prcomp(select(teams, points3, points2, free_throws, off_rebounds, def_rebounds,
                           assists, steals, blocks, turnovers, fouls), scale. = TRUE)
teams_pca
```

**Createa a data frame with the eigenvalues:**

```{r}
eigs <- data.frame(
  "eigenvalues" = round(teams_pca$sdev ^ 2, 4),
  "prop" = round( teams_pca$sdev ^ 2 / sum(teams_pca$sdev ^ 2), 4),
  "cumprop" = round(cumsum(teams_pca$sdev ^ 2 / sum(teams_pca$sdev ^ 2)), 4),
  row.names = 1:10
)
eigs

```

**Use the first two PCs to get a scatterplot of the teams**

```{r}
teams_pca$x <- as.data.frame(teams_pca$x)

ggplot(data = teams_pca$x, aes(x = PC1, y = PC2)) + 
  geom_point() +
  geom_text(aes(label = teams$team), vjust = 1.25) +
  labs(title = "PCA plot (PC1 and PC2)") +
  geom_hline(yintercept = 0, size = 1, color = "grey") +
  geom_vline(xintercept = 0, size = 1, color = "grey")
```

### Index based on PC1

**In order to build an index based on the first PC, you are going to transform PC1. To get a
more meaningful scale, you can rescale the first PC with a new scale ranging from 0 to 100.**

```{r}
s_1 <- 100 * (teams_pca$x$PC1 - min(teams_pca$x$PC1)) / (max(teams_pca$x$PC1) - min(teams_pca$x$PC1))
```

**Once you have obtained the rescaled PC1, you can produce a barchart like the previous ones:**

```{r}
teams$PC1 = s_1
ggplot(data = teams , aes(x = reorder(team, PC1), y = PC1)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  labs(y = "First PC (scaled from 0 to 100)", x = "Team", title = "NBA Teams ranked by scaled PC1")
```

**Provide a brief description of the PC1 index to rank the teams.**\
*Ranked by the first PC, GSW ranked the first way higher than the second with index 100; while SAC ranked the last with index 0.*

# Comments and Reflections

- **Was this your first time working on a project with such file structure? If yes, how do
you feel about it?**\
Yes, it was. The file structure is very organized.\

- **Was this your first time using relative paths? If yes, can you tell why they are important
for reproducibility purposes?**\
Yes, it was. Because it will ensure that if you move the project location, you can still access each file without an error.\

- **Was this your first time using an R script? If yes, what do you think about just writing
code?**\
Yes, it was. I feel more comfortable becasue I don't need to deal with the format of output.\

- **What things were hard, even though you saw them in class/lab?**\
The PCA process and the meaning of doing each step.\

- **What was easy(-ish) even though we haven't done it in class/lab?**\
Using ggplot to generate different plots.\

- **Did anyone help you completing the assignment? If so, who?**\
No. I did everything on my own.\

- **How much time did it take to complete this HW?**\
Around 4 hours.\

- **What was the most time consuming part?**\
Understanding the PCA process.\

- **Was there anything interesting?**\
Using ggplot to generate different plots.\


