---
title: "Lab01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

variables
```{r}
phone <- 80
transportation <- 20
groceries <- 527
gym <- 10
rent <- 1500
other <- 83
```

total expense
```{r}
total <- phone + transportation + groceries + gym + rent + other
total
```

semester expenses
```{r}
total * 5
```

year expenses
```{r}
total * 10
```

your vector expenses
```{r}
expense <- c(phone, transportation, groceries, gym, rent, other)
barplot(expense)
```

```{r}
expense <- sort(expense, decreasing = TRUE)

## Find a range of y's that'll leave sufficient space above the tallest bar
ylim <- c(0, 1.1*max(expense))
## Plot, and store x-coordinates of bars in xx
expense_plot <- barplot(expense, ylim = ylim, names.arg = c("rent", "groceries",                            "other", "phone", "transport", "gym"))
## Add text at top of bars
text(x = expense_plot, y = expense, label = expense, pos = 3, cex = 0.8, col = "red")
```

# Pythagoras formula
```{r}
a <- 3
b <- 4
c <- sqrt(a ^ 2 + b ^ 2)
c
```

# Binomial Formula
Let's consider a simple example. A fair coin is tossed 5 times. What is the probability of getting exactly 2 heads?

Create the objects n, k, and p for the number of trials, the number of success, and the probability of success, respectively.
```{r}
n <- 5
k <- 2
p <- 1/2
```

Use factorial() to compute the number of combinations "n choose k"
```{r}
n_choose_k <- factorial(n)/(factorial(k) * factorial(n - k))
```

Apply the binomial formula, using factorial(), to calculate the probability of getting exactly 2 heads out of 5 tosses.
```{r}
n_choose_k * (p ^ k) * (1 - p) ^ (n - k)
```


Recalculate the same probability but now using choose() (instead of factorial())
```{r}
choose(n, k) * (p ^ k) * (1 - p) ^ (n - k) 
```

Consider rolling a fair die 10 times. What is the probability of getting exactly 3 sixes?
```{r}
n2 <- 10
k2 <- 3
p2 <- 1/6
choose(n2, k2) * (p2 ^ k2) * (1 - p2) ^ (n2 - k2) 
```


Now look for help documentation (e.g. help.search() or ??) using the keyword binomial: binomial.
```{r}
??binomial
```

You should get a list of topics related with the searched term binomial.

Choose the one related with the Binomial Distribution, which is part of the R package stats (i.e. stats::Binomial).

Read the documentation and figure out how to use the dbinom() function to obtain the above probabilities: 2 heads in 5 coin tosses, and 3 sixes in 3 rolls of a die.
```{r}
dbinom(k, n, p)
dbinom(k2, n2, p2)
```


How would you modify the previous binomial function to calculate the same probability (2 heads in 5 tosses) of a biased coin with a chance of heads of 35%?
```{r}
dbinom(k, n, p = 0.35)
```


Finally, obtain the probability of getting more than 3 heads in 5 tosses with a biased coin of 35% chance of heads.
```{r}
dbinom(4, 5, 0.35) + dbinom(5, 5, 0.35)
```
# Installing Packages
Install packages "stringr", "RColorBrewer", and "XML"
```{r}
install.packages(c("stringr", "RColorBrewer", "XML"), repos = "http://cran.us.r-project.org")
```

Calculate: 3x2^2+ 4x+8 when x=2
```{r}
x <- 2
3 * 2 ^ 2 + 4 * x + 8
```

Calculate: 3x2+4x+8 but now with a numeric sequence for x using x <- -3:3
```{r}
x <- -3:3
3 * 2 ^ 2 + 4 * x + 8
```

