---
title: "warmup02"
author: "Xiaoya Li"
date: "September 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("~/UC_Berkeley/Stat133/HW2/nba2017-salary-points.RData")
ls()
```

##Exploratory Data Analysis

<br>

###Quantitative Variable


- Use the summary() function to get a quick summary of descriptive statitstics for each numeric vector
```{r}
summary(points)
```

```{r}
summary(points1)
```

```{r}
summary(points2)
```

```{r}
summary(points3)
```

```{r}
summary(salary)
```

- Now, look up for functions that allow you to get the following statistics:

    * mean (i.e. average)
    * standard deviation
    * minimum value
    * maximum value
    * median
    * quartiles
    
```{r}
mean(points)
sd(points)
min(points)
max(points)
median(points)
quantile(points)
```
    
- What are the typical values in each vector?  
  In general, players get a average `points` of 547, with a standard deviation of 489. The minimum points is 0 while the maximum is 2558. The median of all the points is 432. 


- What's the spread in each vector?  
  The interquantiles of `points` is 0, 156, 432, 780, 2558.
  
  
- Look at the distribution: use hist() and boxplot()

```{r}
hist(points)
boxplot(points, horizontal = TRUE)
```

- You can also try to get a density curve (or density polygon). Find out how to do this.
```{r}

plot(density(points), ylab = "frequency(%)", xlab = "points" )


```


##Qualitative Variable
- Pick one of the categorical variables: `team` or `position`  
- If the variable that you choose is not an R factor, then use factor() to convert the object into a factor.  
- Use table() to get a frequency table (i.e. counts of each category).  
```{r}
team_fre <- table(team)
team_fre
```

- Find out how to use the obtained frequency table to calculate relative frequencies (proportions).

```{r}
team_refre <- team_fre / length(team)
team_refre
```

- Use the frequencies (counts) and relative frequencies (proportions) to describe the overall distribution.
  All the team has either 14 players or 15 players.

- Use barplot() to display the frequencies with a barchart.
```{r}
barplot(team_fre, ylim=c(0,20))
```


##Report your process
- What things were hard, even though you saw them in class?  
  Understand what factor really is.


- What was easy(-ish) even though we haven't done it in class?  
  Using the `??` and `?` to search documentation for functions.
  
  
- What type of "errors" you struggled with (if any)?  
  To generate a more descriptive graph, including some labels and data on the graph. 
  
  
- What are the parts you are not fully understanding?    
  What is factor and what's the different between vector.
  
  
- What was the most time consuming part?  
  Trying to figure out some specific instruction. 
  
- Did you collaborate with other students? If so, with who? In what manner?  
  No. I worked alone. 
  
  
- Was there any frustrating issue? (e.g. RStudio cryptic error, one or more package not playing nice)   Markdown syntax sometimes doesn't work well. 