---
title: "Lab08"
author: "Xiaoya Li"
date: "October 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For loops
rewrite the for loop in order to triple every element in vec, and printing the output at each step of the loop:

```{r}
vec <- c(4, 7, 10) # Change this value!

for (j in 1:3) { # Replace c() with an appropriate sequence.
  # Fill in.
  vec[j] = 3 * vec[j]
  print(vec[j])
}
```

What if you want to create a vector vec2, in which you store the values produced at each iteration of the loop? Here's one possibility:

```{r}
vec <- c(4, 7, 10)  # Change this value!
vec2 <- rep(0, length(vec))  # "empty" of zeros vector to be filled in the loop

for (i in 1:3) {# Replace c() with an appropriate sequence.
  # Fill in.
  vec2[i] = 3 * vec[i]
  print(vec2[i])
}
```

## Summation Series

Write a for loop to compute the following two series. Your loop should start at step k = 0 and stop at step n. Test your code with different values for n. And store each k-th term at each iteration. Does the series converge as n increase?

$$ \sum_{k=0}^{n} \frac{1}{2^k} = 1 + \frac{1}{2} + \frac{1}{4} + \frac{1}{8} + \dots + \frac{1}{2^n} $$

```{r}
summation1 <- function(n){
  vec <- rep(0, n + 1)
  sum = 0
  for (i in 0:n) {
   vec[i + 1] = 1 / (2 ^ i)
   sum = sum + vec[i + 1]
  }
  print(sum)
  return(vec)
}

n_10 = summation1(10)
n_100 = summation1(100)
n_1000 = summation1(1000)
# The serie converges to 2 as n increase.
```

$$ \sum_{k=0}^{n} \frac{1}{9^k} =1 + \frac{1}{9} + \frac{1}{81} + \dots + \frac{1}{9^n} $$

```{r}
summation2 <- function(n){
  vec <- rep(0, n + 1)
  sum = 0
  for (i in 0:n) {
   vec[i + 1] = 1 / (9 ^ i)
   sum = sum + vec[i + 1]
  }
  print(sum)
  return(vec)
}

n_10 = summation2(10)
n_100 = summation2(100)
n_1000 = summation2(1000)
# The serie converges to 1.125 as n increase.
```

## Arithmetic Series

Write a for loop to compute the following arithmetic series an = a1 + (n ??? 1)d when a1 = 3, and d = 3. For instance: 3 + 6 + 12 + 24 + ..
Test your code with different values for n. And store each n-th term at each iteration. Does the series converge as n increase?

```{r}
arithmetic1 <- function(n){
  a_n = rep(0, n)
  a_1 = 3
  d = 3
  sum = 0
  for (i in 1:n){
    a_n[i] = a_1 + (i - 1) * d
    sum = sum + a_n[i]
  }
  print(sum)
  return(a_n)
}

n_10 <- arithmetic1(10)
n_100 <- arithmetic1(100)
n_1000 <- arithmetic1(1000)
#the series does not converge as n increase
```

## Geometric Sequence

Write a for loop to compute the sum of the first n terms of: 3 + 6 + 12 + 24 + .. Test your code with different values for n. Does the series converge as n increase?

```{r}
geometric <- function(n){
  a_n = rep(0, n)
  a_1 = 3
  r = 3
  sum = 0
  for (i in 1:n){
    a_n[i] = a_1 * (r ^ (i - 1))
    sum = sum + a_n[i]
  }
  print(sum)
  return(a_n)
}

n_10 <- geometric(10)
n_100 <- geometric(100)
n_1000 <- geometric(1000)
# the series does not converge as n increase
```

## Sine Approximation

Consider the following series that is used to approximate the function sin(x):
$$ sin(x) \approx x - \frac{x^3}{3!} + \frac{x^5}{5!} - \frac{x^7}{7!} + \dots $$

Write a for loop to approximate sin(x). Try different number of terms, n = 5, 10, 50, 100. Compare your loop with the sin() function.

```{r}
my_sin <- function(x,n){
  sum = 0
  for(i in 1:n){
    sum = sum + (-1) ^ (i + 1) * x ^ (2 * i - 1) / factorial(2 * i - 1)
  }
  print(sum)
}

sin(5)
my_sin(5,5)
my_sin(5,10)
my_sin(5,50)
my_sin(5,100)
```

## For loop with a matrix

Consider the following matrix X:

```{r}
set.seed(123)
X <- matrix(rnorm(12), nrow = 4, ncol = 3)
X
```

Write code in R, using loops, to get a matrix Y such that the negative numbers in A are transformed into squared values, while the positive numbers in A are transformed into square root values

```{r}
Y <- matrix(nrow = 4, ncol = 3)

for(i in 1:4){
  for(j in 1:3){
    x = X[i,j]
    if(x < 0){
      Y[i, j] = x ^ 2
    }
    else{
      Y[i, j] = sqrt(x)
    }
  }
}

Y
```

## Dividing a number by 2 multiple times
## Make a reduce() function

Now generalize the above code to create a function reduce() which performs the same operation. (You should change very little.)

```{r}
# your reduce() function
reduce <- function(x) {
  # Fill in.
  while (x %% 2 == 0) { # Continue the loop as long as val_while is even.
   x <- x / 2
   print(x)
  }
}

reduce(898128000)

```

## Average

The average of n numbers x1, x2, ., xn is given by the following formula:
$$ \bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i = \frac{x_1 + x_2 + \dots + x_n}{n} $$

Write R code, using each type of loop (e.g. for, while, repeat) to implement the arithmetic mean of the vector x = 1:100

```{r}
x = 1:100
sum = 0
for(i in x){
  sum = sum + x[i]
}
average = sum / length(x)
average
mean(x)
```

## Standard Deviation

The sample standard deviation of a list of n numbers x1, x2, ., xn is given by the following formula:
$$ SD = \sqrt{ \frac{1}{n-1} \sum_{i=1}^{n} (x_i - \bar{x})^2 } $$

Write R code, using each type of loop (e.g. for, while, repeat) to implement the sample standard deviation of the vector x = 1:100

```{r}
sum = 0
for(i in x){
  sum = sum + (x[i] - average) ^ 2
}
standard_deviation = sqrt(sum / (length(x) - 1))
standard_deviation
sd(x)

```

## Geometric Mean

The geometric mean of n numbers x1, x2, ., xn is given by the following formula:
$$ \bar{x} = \left ( \prod_{i=1}^{n} x_i \right )^{1/n} $$

Write R code, using each type of loop (e.g. for, while, repeat) to implement the geometric mean of the vector x = 1:50

```{r}
x = 1:50
product = 1
for(i in x){
  product = product * x[i]
}
geo_mean <- product ^ (1 / length(x))
geo_mean
```

