---
title: "Post02: Advanced Visualization"
author: "Xiaoya Li"
date: "November 26, 2017"
output: github_document
always_allow_html: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


## Introduction

<br>

Data visulization is one of the most powerful usage of R. So far, we've seen many 
type of data visualization, for example, Histogram, Barchart, Boxplot, Scatter 
plot, etc. These are common and efficient, but other than that, there are still 
many choices of interesting data visualization. In today's post, I'm going to 
introduce three of them: Heat Map, Mosaic Plot and 3D Graph.

<br>
<br>

## Advanced Visualization

Before we jump into the topic, let's load all packages we are going to use in 
this post.

```{r}
#datasets contains most of the database we are going to use
library(datasets)
library(boot)

#graphic
library(ggplot2)
library(ggmosaic)
library(plotly)
```

<br>
<br>


### I. Heat Map


Heat map represents relationship between two or more variable in a two 
demensional graph, using intensity of color in a tabular format.


<br>
 
**Example 1**


<br>

For the first example, let's see how to use the `heatmap()` function in R to 
generate a heat map.\
The `USJudgeRatings` dataset contains numeric rating of selected judges from 
different persepective. Let's take a look.\

```{r}
head(USJudgeRatings)
```

`heatmap()` function requires our input to be a matrix. 

```{r}
#convert it to a matrix
heatmap(as.matrix(USJudgeRatings))
#add a title
title(main = "Heat Map of Judge Rating")
```

*Description: This graph represents the rating of each judge. For each column, 
the dark portion indicates the best judge in this attribute; for each row, the 
dark portion indicates the best rating of each judge got.*


================================================================================

<br>
<br>

**Example 2**


<br>

Alternatively, we can also use `ggplot` to plot the heat map.
`airquality` dataset contains air quality measurement in New York City.

```{r}
head(airquality)
```

This time, other than plotting all the attribute as the previous graph, we want
to compare the temperature on each day in each month. 
Here's the simple R code using `ggplot()` and `geom_tile()`.

```{r}
ggplot(airquality, aes(Day, Month)) +
  #build the heat map
  geom_tile(aes(fill = Temp)) +
  
  #reverse legend color such that dark color related to higher temperature 
  scale_fill_gradient(high = "#132B43", low = "#56B1F7") +
  
  #add a title
  ggtitle(label = "Heat Map") 
  
```

*Description: This graph represents the temperature of each day in each month. 
Horizontally, the dark portion indicates the hottest day in this month; vertically, 
the dark portion indicates the hottest month.*


================================================================================

<br>
<br>
 
### Mosaic Plot

Mosaic plot is useful to represents categorical data.\


<br>

**Example 3**

<br>

`mosaicplot()` function in R provides a simple way to generate a mosaic plot.\
`HairEyeColor` dataset contains the Hair and Eye Color of Statistics Students.\ 
Let's take a look at the dataset.

```{r}
HairEyeColor
```

`HairEyeColor` is in the class of `table`, which is different from what we used 
to have in data.frame.\


Here's the one line simple R code.

```{r}
mosaicplot(HairEyeColor)
```

*Description: This graph represents the proportion of students of different hair
and eye colors in the size of the area.*


================================================================================

 <br>
  <br>

**Example 4**


 <br>

Still, we can use `ggplot()` to graph a detailed mosaic plot. Inorder to graph
in ggplot, we need to add another package `ggmosaic`.\


`Titanic` is a bit more complicated dataset, it contains information of Survival
of passengers on the Titanic. Also, the `Titanic` in `dastasets` is in the class 
of table, but since `ggplot()` works with dataframe instead of table, we need the 
`Titanic` data in its original form. 

```{r}
#download .csv file online
download.file(url = "https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/master/csv/datasets/Titanic.csv", destfile = "../Data/titanic_raw.csv")

#read in data
titanic_raw <- read.csv(file = "../Data/titanic_raw.csv")
head(titanic_raw)
```

We need some data preparation before we draw the graph.\
We add a column `AgeDecade` that categorize each passengers' age into decade. 

```{r}
#Initialize a empty character vector
AgeDecade <- rep(" ", nrow(titanic_raw))

#Use a for loop to categorize each passengers' age
for(i in 1:nrow(titanic_raw)){
  x = titanic_raw$Age[i]
  if(is.na(x)){AgeDecade[i] = "unknown"}
  else if(x >= 0 & x < 10){AgeDecade[i] = "0 - 9"}
  else if(x >= 10 & x < 20){AgeDecade[i] = "10 - 19"}
  else if(x >= 20 & x < 30){AgeDecade[i] = "20 - 29"}
  else if(x >= 30 & x < 40){AgeDecade[i] = "30 - 39"}
  else if(x >= 40 & x < 50){AgeDecade[i] = "40 - 49"}
  else if(x >= 50 & x < 60){AgeDecade[i] = "50 - 59"}
  else if(x >= 60 & x < 70){AgeDecade[i] = "60 - 69"}
  else{AgeDecade[i] = "70 +"}
}

#Add that vector into a column of the dataframe
titanic_raw$AgeDecade <- as.factor(AgeDecade)
head(titanic_raw)
```

Now, we are ready to graph. I want to visualize the proportion of survivals within
sex and age interval.

```{r}
ggplot(titanic_raw) +
  #pick the attribute we are interested in
  geom_mosaic(aes(x = product(Survived,AgeDecade), fill = factor(Survived))) +
  
  #adjust the x-axis to make it more readable
  theme(axis.text.x = element_text(angle = -25, hjust = .1)) +
  
  #facet the data into female and male
  facet_grid(Sex ~ .) +
  
  #add title of the graph
  labs(x="Age in Decades ", title='f(Survived, AgeDecade | Gender)') +
  
  #add label to legend
  scale_fill_discrete(name="Survived", labels = c("No", "Yes")) 
```

*Description: This graph visually shows the proportion of survivals within 
different age intervals and sex. Generally, more female survived than male. And 
people in age 20-29 survived most in both female and male.*


================================================================================

 <br>
  <br>

### III. 3D Graph


Obviously, the 3D graph is one of the impressive graphs of all, it can represent 
relation between up tp three variables visually straightforward.\


To draw a 3D graph, package `plotly` is a nice tool combined with `ggplot`.\

 <br>

**Example 5**

 <br>

For this example, I'm showing you how to graph the most basic 3d scatterplot.\


The dataset we used is `cane`, which contains the Sugar-cane Disease Data. Since 
the column name are not meaningful enough, here's the documentation from R about 
this dataset.


> n: The total number of shoots in each plot.\
> r: The number of diseased shoots.\
> x: The number of pieces of the stems, out of 50, planted in each plot.\
> var: A factor indicating the variety of sugar-cane in each plot.\
> block: A factor for the blocks.\


Let's take a look of the data.

```{r}
head(cane)
```

We're going to visualize the relation of total number of shoots, number of 
diseased shoots and number of pieces of the stems. Here's the R code:

```{r}
plot_ly(cane, x = ~n, y = ~r, z = ~x, color = ~block,
        colors = c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070')) %>% 
  
  #add label to axis
  add_markers() %>% 
  
  #change the label of the axis
  layout(scene = list(
    xaxis = list(title = '# shoots in each plot'),
    yaxis = list(title = '# diseased shoots'),
    zaxis = list(title = '# pieces of the stems')),
    
    #change the label of the legend
         annotations = list(
           x = 1.13, y = 1.05,
           text = 'Block',
           xref = 'paper',
           yref = 'paper',
           showarrow = FALSE))
  
```

*Note: You can interact with the graph by clicking the legend.*\

*Description: This 3d scatter plot shows relation of total number of shoots, 
number of diseased shoots and number of pieces of the stems, with different color
representing different blocks.*


================================================================================

 <br>
  <br>

**Example 6**

 <br>

In this example, I'm showing to how to use 3D surface graph to visualize the 
probability density estimation.\


`cars` in packages `dataset` contains infomation of Speed and Stopping Distances 
of Cars. 

```{r}
head(cars)
```

What if we want to estimate the stopping distance given the speed of the car?
**Kernel density estimation** is a way to estimate the probability density 
function of a random variable. To learn more about **Kernel density estimation**,
[click here](https://en.wikipedia.org/wiki/Kernel_density_estimation).\


There's a function in package `MASS` called `kde2d()`, "Two-Dimensional Kernel 
Density Estimation", that allows us to carry out the estimation.\


You would end up with a list with x being your independent variable and y being 
the dependent variable that you're trying to estimate based on x.\

```{r}
#prepare the data using kernel density estimation

#MASS:: allows us to call functions without actually loading MASS package

#n is the number of grid points in each direction, 
#the larger of n, the smoother your graph would looks like
kd <- with(cars, MASS::kde2d(x = speed, y = dist, n = 50))

str(kd)
```


```{r}
plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% 
  
  #plot the 3d surface 
  add_surface() %>% 
  
   #change the label of the axis
  layout(scene = list(
    xaxis = list(title = 'Speed (mph)'),
    yaxis = list(title = 'Stopping Distance (ft)')))
```

*Description: This 3D surface estimate the Stopping distance of a car giving 
its speed. Rotate the object to see more detail.*


================================================================================

 <br>
  <br>
  
  
## Conclusion

I hope this post can give you more idea about data visulization.

- To visualize data from many perspectives at the same time, consider **Heat Map**.

- To visualize categorical data, try **Mosaic Plot**.

- To visualize multi-variables, try using different kind of **3D Graph**.
 
 
  <br>
   <br>
   
   
## Reference 


1. http://flowingdata.com/2010/01/21/how-to-make-a-heatmap-a-quick-and-easy-solution/
2. https://www.r-bloggers.com/7-visualizations-you-should-learn-in-r/
3. https://www.analyticsvidhya.com/blog/2015/07/guide-data-visualization-r/
4. https://www.statmethods.net/advgraphs/mosaic.html
5. https://cran.r-project.org/web/packages/ggmosaic/vignettes/ggmosaic.html
6. https://plot.ly/r/3d-scatter-plots/#basic-3d-scatter-plot
7. https://plot.ly/r/3d-surface-plots/